Model Search in Relational Domains

### 2016-10-22 deductions.tex

%-------------------------------------------------------------------------------
\subsection{Abstract Deduction Transformers} \label{sec:abst}
%-------------------------------------------------------------------------------
\Omit{
To make our algorithm efficient, we have to focus the abstract
transformers on performing only the minimally necessary
work. 
}
\rmcmt{The key considerations for building an abstract transformer are 
precision and efficiency.  An precise transformer is usually less efficient, 
and vice-versa.  We provide a parameterised abstract transformer that allows us to 
control precision and efficiency through propagation heuristics. }
%
We thus define a specialised variant of the abstract transformer to compute
deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$,
which we call \emph{abstract deduction transformer}.
%
A subdomain contains a chosen subset of the elements in $\domain$ including
$\bot$ and $\top$, \rmcmt{which forms a lattice}.  An abstract 
deduction transformer is formally defined as follows. 
\begin{equation}\label{eq:at2}
\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\in\gamma_\domain(\absval), \val\models \constraint\})
\end{equation}
For $\subdomain=\domain$, the abstract deduction transformer is identical 
to the abstract transformer defined in Eq.~(\ref{eq:abstrans})
in Section~\ref{sec:domains}.  We denote $\abstrans{\domain}{\constraint}^\subdomain$
as $\abstransel{\subdomain}$.

@@

%The first advantage of this definition is that 
\rmcmt{
This parameterisation of the abstract deduction transformer with a subdomain 
$\subdomain$ allow us to guide the propagation in {\em forward}, {\em backward} 
or {\em multi-way} direction.
}
%using our abstract transformers.
%we can control how we perform 
%
For example, for $\absval=(0\leq y \leq 1 \wedge 5\leq z)$, we have
$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{x\}]}(\absval)=\absval\meet(x\geq
6)$. For $x=y+z$, this performs a right-hand side to left-hand side propagation and
hence emulates a forward analysis, whereas
$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{y,z\}]}=\absval\meet\top$
emulates a backward analysis.

We call the propagation with arbitrary subdomains,
e.g. $\subdomain=\domain$, \emph{multi-way propagation}, which is able
to simultaneously perform forward and backward propagation.
%
\Omit{
Note that a restriction to a subdomain makes a transformer less
precise.
}
Note that a restricted subdomain makes a transformer less precise, but more
efficient.  On the contrary, an unrestricted subdomain make a transformer more 
precise, but less efficient. Therefore, we have the property
$\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq
\abstrans{\domain}{\constraint}^\subdomain(\absval)$.
%However, the choice of a subdomain makes the deductions more efficient.
\rmcmt{Thus, the choice of a subdomain helps to fine-tune the deductions 
during the propagation phase.}

\paragraph {\textbf{Lazy closure computation}} 
Computing closure for relational domains, such as octagons is extremely
expensive~\cite{pldi15}.  An advantage of our formalism in 
Eq.~(\ref{eq:at2}) is that
%we can compute the
the \emph{closure} operation for relational domains can be computed 
in a lazy manner. To this end, we construct a subdomain 
$\subdomain=\makesubdomain_\domain(\subvars)$ for $\abstransel{\subdomain}$,
which is sufficient to perform one step of the closure operation when 
$\abstransel{\subdomain}$ is applied.
%
For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
$\subvars=\{y\}$. An octagonal inequality relates at 
most two variables. Thus it is sufficient to consider the subdomain
$\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
which will compute the one-step transitive relations of~$y$ with each
of the other variables. 
%
\rmcmt{Only if a abstract deduction transformer subsequently makes new deductions 
on $x$ or $z$, then the next step of the closure will be computed through 
the subdomain $\octagons[\{x,z\}]$.}
\Omit{
Hence, when the abstract deduction transformer is applied we do not
compute the full closure in the full domain,
but we compute only a single step of the closure in a restricted
domain, which makes single deduction steps more efficient.
}
Hence, an application of abstract deduction transformer does not 
compute the full closure in the full domain, but compute only a 
single step of the closure in a restricted domain, which makes each 
deduction step more efficient.  Thus, we delay the closure operation 
until the point where it is absolutely necessary. 
%
\Omit{
If $\abstransel{\subdomain}$ deduces new information about $x$ or $z$
then the next step of the closure will be computed by the worklist
mechanism of Algorithm~\ref{Alg:ms} that we describe next.
}
%

>>>>

%-------------------------------------------------------------------------------
\subsection{Abstract Deduction Transformers} \label{sec:abst}
%-------------------------------------------------------------------------------
\Omit{
To make our algorithm efficient, we have to focus the abstract
transformers on performing only the minimally necessary
work. 
}
The key considerations for building an abstract transformer are 
precision and efficiency.  An precise transformer is usually less efficient, 
and vice-versa.  We provide a parameterised abstract transformer that allows us to 
control precision and efficiency through propagation heuristics. 
%
We thus define a specialised variant of the abstract transformer to compute
deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$,
which we call \emph{abstract deduction transformer}.
%
A subdomain contains a chosen subset of the elements in $\domain$ including
$\bot$ and $\top$, which forms a lattice.  An abstract 
deduction transformer is formally defined as follows. 
\begin{equation}\label{eq:at2}
\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\in\gamma_\domain(\absval), \val\models \constraint\})
\end{equation}
For $\subdomain=\domain$, the abstract deduction transformer is identical 
to the abstract transformer defined in Eq.~(\ref{eq:abstrans})
in Section~\ref{sec:domains}.  We denote $\abstrans{\domain}{\constraint}^\subdomain$
as $\abstransel{\subdomain}$.

@@

%The first advantage of this definition is that 
This parameterisation of the abstract deduction transformer with a subdomain 
$\subdomain$ allow us to guide the propagation in {\em forward}, {\em backward} 
or {\em multi-way} direction.
%using our abstract transformers.
%we can control how we perform 
%
For example, for $\absval=(0\leq y \leq 1 \wedge 5\leq z)$, we have
$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{x\}]}(\absval)=\absval\meet(x\geq
6)$. For $x=y+z$, this performs a right-hand side to left-hand side propagation and
hence emulates a forward analysis, whereas
$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{y,z\}]}=\absval\meet\top$
emulates a backward analysis.

We call the propagation with arbitrary subdomains,
e.g. $\subdomain=\domain$, \emph{multi-way propagation}, which is able
to simultaneously perform forward and backward propagation.
%
\Omit{
Note that a restriction to a subdomain makes a transformer less
precise.
}
Note that a restricted subdomain makes a transformer less precise, but more
efficient.  On the contrary, an unrestricted subdomain make a transformer more 
precise, but less efficient. Therefore, we have the property
$\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq
\abstrans{\domain}{\constraint}^\subdomain(\absval)$.
%However, the choice of a subdomain makes the deductions more efficient.
Thus, the choice of a subdomain helps to fine-tune the deductions 
during the propagation phase.

\paragraph {\textbf{Lazy closure computation}} 
Computing closure for relational domains, such as octagons is extremely
expensive~\cite{pldi15}.  An advantage of our formalism in 
Eq.~(\ref{eq:at2}) is that
%we can compute the
the \emph{closure} operation for relational domains can be computed 
in a lazy manner. To this end, we construct a subdomain 
$\subdomain=\makesubdomain_\domain(\subvars)$ for $\abstransel{\subdomain}$,
which is sufficient to perform one step of the closure operation when 
$\abstransel{\subdomain}$ is applied.
%
For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
$\subvars=\{y\}$. An octagonal inequality relates at 
most two variables. Thus it is sufficient to consider the subdomain
$\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
which will compute the one-step transitive relations of~$y$ with each
of the other variables. 
%
Only if a abstract deduction transformer subsequently makes new deductions 
on $x$ or $z$, then the next step of the closure will be computed through 
the subdomain $\octagons[\{x,z\}]$.
\Omit{
Hence, when the abstract deduction transformer is applied we do not
compute the full closure in the full domain,
but we compute only a single step of the closure in a restricted
domain, which makes single deduction steps more efficient.
}
Hence, an application of abstract deduction transformer does not 
compute the full closure in the full domain, but compute only a 
single step of the closure in a restricted domain, which makes each 
deduction step more efficient.  Thus, we delay the closure operation 
until the point where it is absolutely necessary. 
%
\Omit{
If $\abstransel{\subdomain}$ deduces new information about $x$ or $z$
then the next step of the closure will be computed by the worklist
mechanism of Algorithm~\ref{Alg:ms} that we describe next.
}
%
